---
title: "Course Project"
author: "Weslley Moura"
date: "Sunday, October 30, 2016"
output: html_document
---

## Resume

**Human Activity Recognition**

Human Activity Recognition - HAR - has emerged as a key research area in the last years and is gaining increasing attention by the pervasive computing research community (see picture below, that illustrates the increasing number of publications in HAR with wearable accelerometers), especially for the development of context-aware systems. There are many potential applications for HAR, like: elderly monitoring, life log systems for monitoring energy expenditure and for supporting weight-loss programs, and digital assistants for weight lifting exercises.

[Source: http://groupware.les.inf.puc-rio.br/](Source: http://groupware.les.inf.puc-rio.br/)

**This project**
In this project we are going to identify the quality of barbell bicep curls by using data from belt, forearm, arm, and dumbbell monitors.

## Libraries
```{r}
library(caret)
library(caretEnsemble)
library(rpart)
library (randomForest)
library(fastAdaboost)
```

## Reproducible code
```{r}
set.seed(3007)
```

## Loading datasets
```{r}
setwd("C://MeusArquivos//datasciencecoursera//PracticalML")
dataset.train.all<-read.table("./pml-training.csv", header=TRUE, sep=",")
dataset.test<-read.table("./pml-testing.csv",header=TRUE, sep=",")
```

## Feature selection
Source: Idea extrated from https://rpubs.com/kdalve/pmachlearn

Training dataset
```{r}
trainingaccel<-grepl("^accel",names(dataset.train.all))
trainingtotal<-grepl("^total",names(dataset.train.all))
roll<-grepl("^roll",names(dataset.train.all))
pitch<-grepl("^pitch",names(dataset.train.all))
yaw<-grepl("^yaw",names(dataset.train.all))
magnet<-grepl("^magnet",names(dataset.train.all))
gyro<-grepl("^gyro",names(dataset.train.all))
acceldata<-dataset.train.all[ ,trainingaccel]
rolldata<-dataset.train.all[ ,roll]
pitchdata<-dataset.train.all[ ,pitch]
yawdata<-dataset.train.all[,yaw]
magnetdata<-dataset.train.all[,magnet]
gyrodata<-dataset.train.all[,gyro]
totaldata<-dataset.train.all[,trainingtotal]
dataset.train.all<-cbind(acceldata,rolldata,pitchdata,yawdata,magnetdata,gyrodata,totaldata,dataset.train.all[ ,160])
colnames(dataset.train.all)[53]<-'target'
```

Testing dataset
```{r}
testingaccel<-grepl("^accel",names(dataset.test))
testingtotal<-grepl("^total",names(dataset.test))
troll<-grepl("^roll",names(dataset.test))
tpitch<-grepl("^pitch",names(dataset.test))
tyaw<-grepl("^yaw",names(dataset.test))
tmagnet<-grepl("^magnet",names(dataset.test))
tgyro<-grepl("^gyro",names(dataset.test))
tacceldata<-dataset.test[ ,testingaccel]
trolldata<-dataset.test[ ,troll]
tpitchdata<-dataset.test[,tpitch]
tyawdata<-dataset.test[,tyaw]
tmagnetdata<-dataset.test[,tmagnet]
tgyrodata<-dataset.test[,tgyro]
ttotaldata<-dataset.test[,testingtotal]
dataset.test<-cbind(tacceldata,trolldata,tpitchdata,tyawdata,tmagnetdata,tgyrodata,ttotaldata,dataset.test[ ,160])
colnames(dataset.test)[53]<-'id'
```

## Splittig trainig dataset
```{r}
trainIndex <- createDataPartition(dataset.train.all$target, p=0.80, list=FALSE)
dataset.train <- dataset.train.all[ trainIndex,]
dataset.validation <- dataset.train.all[-trainIndex,]
```

## Model selection

In the train control I am using a k-fold cross validation.

```{r}
# Train control
trainControl <- trainControl(method="cv", number=10, classProbs=TRUE, savePredictions=TRUE, index=createFolds(dataset.train$target))

# Algorithms
algorithmList <- c('rpart', 'rf')

# Create models
models <- caretList(target~., data=dataset.train, trControl=trainControl, methodList=algorithmList)

# Ckeck results
results <- resamples(models)
```

## Comparing results
```{r}
# Compare results for each model
summary(results)

# Using a box plot to compare results
scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(results, scales=scales)
```

## Baseline model

The best model was selected as the baseline.
```{r}
fit.rf <- train(target~., data=dataset.train, method="rf", trControl=trainControl)
```

## Making predictions on validation dataset 
```{r}
predictions = predict(fit.rf,dataset.validation)
confusionMatrix(dataset.validation$target,predictions)
```

## Making predictions on test dataset 
```{r}
predictionsTest = predict(fit.rf,dataset.test)
```

## Showing predictions 
```{r}
finalResults = cbind(id = dataset.test$id, target = predictionsTest)
print (finalResults)
```

